Calculating Entropies With Nested Sampling

### 2017-02-28 paper/ms.tex
More writing

\documentclass[entropy,article,submit,oneauthor,pdftex,10pt,a4paper]{mdpi}
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, administrativesciences, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, appliedsciences, arts, atmosphere, atoms, axioms, batteries, behavioralsciences, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsciences, buildings, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, cosmetics, crystals, data, dentistryjournal, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, inorganics, insects, ijerph, ijfs, ijms, ijns, ijgi, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, joi, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, lubricants, machines, marinedrugs, materials, mathematics, medicalsciences, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, nanomaterials, ncrna, nutrients, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, sinusitis, socialsciences, societies, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, universe, vaccines, veterinarysciences, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but could be replaced by using one of the class options:
% article, review, communication, commentary, bookreview, correction, addendum, editorial, changes, supfile, casereport, comment, conceptpaper, conferencereport, meetingreport, discussion, essay, letter, newbookreceived, opinion, projectreport, reply, retraction, shortnote, technicalnote, creative, datadescriptor (for journal Data), briefreport, hypothesis, interestingimage
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
% Please insert a blank line is before and after all equation and eqnarray environments to ensure proper line numbering when option submit is chosen
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option "pdftex" is for use with pdfLaTeX only. If eps figure are used, use the optioin "dvipdfm", with LaTeX and dvi2pdf only.

@@

Bayesian inference is the use of probability theory to
describe uncertainty, often about unknown quantities
(``parameters'') $\x$. Some data $\d$, initially unknown but
thought to be relevant to $\x$, is obtained.
The prior information leads the
user to specify a prior distribution $p(\x)$ for the unknown parameters,
along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
of how the data is related to the parameters
(i.e., if the parameters where known, what data would likely be observed?). By the product rule, this yields a {\em joint prior}
\begin{align}
p(\x, \d) &= p(\x)p(\d | \x)
\end{align}
which is the starting point for inference \citep{caticha}.

The entropy of this joint prior describes the degree to which the question
``what is the value of the pair $(\x, \d)$?'' remains unanswered.

@@

\subsection{Mutual information}
The mutual information is another way of describing the relevance of the
data to the parameters. Its definition, and relation to other quantities, is
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
                       \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
           &= H_{\x} + H_{\d} - H_{\x, \d}\\
           &= H_{\d} - H_{\d | \x}\\
           &= H_{\x} - H_{\x | \d}.
\end{align}
For present purposes, it is more conveniently written as
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
              \log\left[\frac{p(\d | \x)}{p(\d)}\right].
\end{align}
As with the conditional entropy, we see that the difficulty appears in the
form of the log of the marginal likelihood,
$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.

@@

In a Bayesian inference problem with parameters $\x$ and data $\d$, the
mutual information is
\begin{align}
I_{\x, \d} &= \iint p(\x, \d)
                        \ln \left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]
                        \, d\x \, d\d
\end{align}
and this measures the dependence between $\x$ and $\d$. A Monte Carlo
strategy to evaluate this is to sample from $p(\x, \d)$ and average
the value of the logarithm, which involves computing the log-evidence
\begin{align}
\ln p(\d) = \ln\left[\int p(\x)p(\d | \x) \, d\x\right]
\end{align}
for each possible data set. Computing $p(\d)$ for even a single
data set has historically been considered difficult.

However, if we want the mutual information between the data and
{\em a subset of the parameters} (i.e., there are nuisance parameters
we don't care about), things become even more tricky. Suppose we
partition the parameters $\x$ into important parameters
$\bphi$ and nuisance parameters $\boldeta$, such that we want to calculate
the mutual information $I_{\bphi, \d}$. This is given by
\begin{align}
I_{\bphi, \d} &= \iint p(\bphi, \d)
                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
                        \, d\bphi \, d\d
\end{align}

>>>>

\documentclass[entropy,article,accept,oneauthor,pdftex,10pt,a4paper]{mdpi}
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, administrativesciences, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, appliedsciences, arts, atmosphere, atoms, axioms, batteries, behavioralsciences, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsciences, buildings, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, cosmetics, crystals, data, dentistryjournal, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, inorganics, insects, ijerph, ijfs, ijms, ijns, ijgi, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, joi, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, lubricants, machines, marinedrugs, materials, mathematics, medicalsciences, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, nanomaterials, ncrna, nutrients, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, sinusitis, socialsciences, societies, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, universe, vaccines, veterinarysciences, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but could be replaced by using one of the class options:
% article, review, communication, commentary, bookreview, correction, addendum, editorial, changes, supfile, casereport, comment, conceptpaper, conferencereport, meetingreport, discussion, essay, letter, newbookreceived, opinion, projectreport, reply, retraction, shortnote, technicalnote, creative, datadescriptor (for journal Data), briefreport, hypothesis, interestingimage
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
% Please insert a blank line is before and after all equation and eqnarray environments to ensure proper line numbering when option submit is chosen
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option "pdftex" is for use with pdfLaTeX only. If eps figure are used, use the optioin "dvipdfm", with LaTeX and dvi2pdf only.

@@

Bayesian inference is the use of probability theory to
describe uncertainty, often about unknown quantities
(``parameters'') $\x$. Some data $\d$, initially unknown but
thought to be relevant to $\x$, is obtained.
The prior information leads the
user to specify a prior distribution $p(\x)$ for the unknown parameters,
along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
of how the data is related to the parameters
(i.e., if the parameters where known, what data would likely be observed?). By the product rule, this yields a {\em joint prior}
\begin{align}
p(\x, \d) &= p(\x)p(\d | \x)
\end{align}
which is the starting point for inference \citep{caticha2008lectures}.

The entropy of the prior $p(\x)$ describes the degree to which the question
``what is the value of $\x$, precisely?'', while the
entropy of the joint prior $p(\x, \d)$
describes the degree to which the question
``what is the value of the pair $(\x, \d)$, precisely?'' remains unanswered.

@@

\subsection{Mutual information}
The mutual information is another way of describing the relevance of the
data to the parameters. Its definition, and relation to other quantities, is
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
                       \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
           &= H_{\x} + H_{\d} - H_{\x, \d}\\
           &= H_{\d} - H_{\d | \x}\\
           &= H_{\x} - H_{\x | \d}.
\end{align}
It can also be written in terms of the prior, likelihood, and evidence,
as
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x)p(\d | \x)
              \left[\log p(\d | \x) - \log p(\d)\right],
\end{align}
i.e., the mutual information is the prior expected value of the
log likelihood minus the log evidence.
As with the conditional entropy, we see that the difficulty appears in the
form of the log of the marginal likelihood,
$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.

@@

In a Bayesian inference problem with parameters $\x$ and data $\d$, the
mutual information is
\begin{align}
I_{\x, \d} &= \iint p(\x, \d)
                        \ln \left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]
                        \, d\x \, d\d \label{eqn:mutual_info2}
\end{align}
and this measures the dependence between $\x$ and $\d$. A
Monte Carlo
strategy to evaluate this is to sample from $p(\x, \d)$ and average
the value of the logarithm, which involves computing the log-evidence
\begin{align}
\ln p(\d) = \ln\left[\int p(\x)p(\d | \x) \, d\x\right]
\end{align}
for each possible data set.
Computing $p(\d)$ for even a single
data set has historically been considered difficult.

However, if we want the mutual information between the data and
{\em a subset of the parameters} (i.e., there are nuisance parameters
we don't care about), things become even more tricky. Suppose we
partition the parameters $\x$ into important parameters
$\bphi$ and nuisance parameters $\boldeta$, such that we want to calculate
the mutual information $I_{\bphi, \d}$. This is given by
\begin{align}
I_{\bphi, \d} &= \iint p(\bphi, \d)
                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
                        \, d\bphi \, d\d \label{eqn:mutual_info3}
\end{align}
which is equivalent to Equation~\ref{eqn:mutual_info2} but with
$\x$ replaced throughout by $\bphi$. However, this is an even more
difficult calculation than before, as $p(\d | \bphi)$, the marginal likelihood
function with the nuisance parameters integrated out, is typically
unavailable in closed form. If we were to marginalise out the nuisance
parameters $\boldeta$ explicitly, this would give us
Equation~\ref{eqn:mutual_info3} with every probability distribution
written as an explicit integral over $\boldeta$:

\begin{align}
I_{\bphi, \d} &= \iint
                   \int p(\bphi, \boldeta)p(\d | \phi, \boldeta) \, d\boldeta
                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
                        \, d\bphi \, d\d \label{eqn:mutual_info3}
\end{align}
